Local PCA Shows How the Effect of Population Structure Differs Along the Genome

### 2016-09-29 writeup/local_pca_paper_with_titlepage.tex
title page

\title{Local PCA Shows How the Effect of Population Structure Differs Along the Genome}
\author{Han Li, Peter Ralph}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
%\section{}
%\subsection{}
% \doublespacing

>>>>

\begin{document}
\vspace*{0.35in}

% Title must be 250 characters or less.
% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
\begin{flushleft}
{\Large
\textbf\newline{Local PCA Shows How the Effect of Population Structure Differs Along the Genome}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Han Li\textsuperscript{1},
Peter Ralph\textsuperscript{2,*}
\\
\bigskip
\bf{1} Molecular and Computational Biology Program, Departments of Biological Sciences, University of Southern California, Los Angeles, CA , USA
\\
\bf{2} Department of Mathematics, University of Oregon, Eugene, OR, USA
\\

\bigskip

* plr@uoregon.edu

\end{flushleft}

### 2016-09-29 writeup/submission/cover-letter.tex
title page

We are writing to submit our paper,
``Local PCA Shows How the Effect of Population Structure Differs Along the Genome'',
for consideration in \textit{JOURNAL XXX}.
In this paper, we set out to describe how background patterns of relatedness 
(colloquially, ``population structure'')
vary along the genome at a scale larger than correlations induced by linkage.
It was not certain that we would find significant large-scale heterogeneity in patterns of relatedness;
in the field we tend to think of neutral, background population structure as a single thing,
uniform across the genome except at a few isolated loci under selection.
What we found was not only substantial variation at chromosomal scales,
but different underlying causes in the three species we consider.
The focus of our paper is on the method we developed to identify and visualize such variation in patterns of relatedness.
We believe our method represents a fundamental advance because it lets the data speak for itself,
providing an unbiased view of what are the most important variation in patterns of relatedness,
rather than looking for specific signals.
The method is therefore complementary to more common methods that identify small regions of the genome
that are outliers for particular summaries of relatedness (e.g., $F_{ST}$-based scans for selection).
We think the paper will be of interest to readers of \textit{JOURNAL XXX}
because the method fills an important gap in the genomicist's toolkit,
that is tailored to describing variation in modern, spatially sampled datasets,
and because it sheds light on the roles of linked selection and chromosomal inversions in 
establishing modern genetic diversity.
% Our results are consistent with what is known about linked selection and chromosomal inversions
% (a possible suprise being no evidence for segregating inversions in \textit{Medicago truncatula});
% the strength of the biological results lie in the unified framework that allows 
% flexible discovery and comparison between the three species.

@@

>>>>

We are writing to submit our paper,
``Local PCA Shows How the Effect of Population Structure Differs Along the Genome'',
for consideration in \textit{PLOS Genetics}.
In this paper, we set out to describe how background patterns of relatedness 
(colloquially, ``population structure'')
vary along the genome at a scale larger than correlations induced by linkage.
It was not certain that we would find significant large-scale heterogeneity in patterns of relatedness;
in the field we tend to think of neutral, background population structure as a single thing,
uniform across the genome except at a few isolated loci under selection.
What we found was not only substantial variation at chromosomal scales,
but different underlying causes in the three species we consider.
The focus of our paper is on the method we developed to identify and visualize such variation in patterns of relatedness.
We believe our method represents a fundamental advance because it lets the data speak for itself,
providing an unbiased view of what are the most important variation in patterns of relatedness,
rather than looking for specific signals.
The method is therefore complementary to more common methods that identify small regions of the genome
that are outliers for particular summaries of relatedness (e.g., $F_{ST}$-based scans for selection).
We think the paper will be of interest to readers of \textit{PLOS Genetics}
because the method fills an important gap in the genomicist's toolkit,
that is tailored to describing variation in modern, spatially sampled datasets,
and because it sheds light on the roles of linked selection and chromosomal inversions in 
establishing modern genetic diversity.
% Our results are consistent with what is known about linked selection and chromosomal inversions
% (a possible suprise being no evidence for segregating inversions in \textit{Medicago truncatula});
% the strength of the biological results lie in the unified framework that allows 
% flexible discovery and comparison between the three species.

@@

If the reviewers would like to read a version of the paper that is single-spaced and has figures included, 
they can download it from \url{http://biorxiv.org/content/early/2016/08/21/070615}.
